A Comparison of Parallel Graph Processing Platforms

### 2017-01-28 papers/gabb17/gabb17-pollard.tex
Added more plots and plot-generation R code.
Also updated some littel bits fo the writeup and some more TODOS.
Makefile now points to the graphics directory but I didn't add any
graphics to the repo yet since they most certainly will change.
Removed execute permission on IEEtran.cls.

\section{Algorithms}
The canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}. The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation. The primary drawback with the Graph500 is it measures a single algorithm: breadth first search (BFS).

This report attempts to add similar rigor to other graph algorithms. This is done by borrowing heavily from the Graph500 specification; we use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19,
C = 0.19,$ and $D = 1-(A+B+C) = 0.05$. The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}. Some clarifications are in order: for the first kernel, the time to read the graph from disk is not considered. Furthermore, these data structures need only be created once and BFS is performed on 64 roots. The second kernel creates a tree with distances from the root, but need only label the existing data structure (and not output the entire graph, for example).

@@

\section{The Existing Systems}
This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}. GraphMat is also used because it has been cited as the most performant [cite]. Our target is shared memory CPU processing.

Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
are not considered here because they are a PITA to get working and the authors never responded to my emails.

\section{Machine Specifications}
Table~\ref{tab:specs} shows the specifications of the research computer (named Arya).

@@

\section{Performance}

@@

\begin{table}[!htb]
	\centering
	\begin{tabular}{l|c|c|c}
		System & Load Graph & Make Data Structure & Run BFS \\ \hline
		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
		GraphBIG & 37.22  & N/A    & 0.1528 \\
		GraphMat & 0.1511 & 1.101  & 0.1031 \\
		GAP      & 2.351  & 0.1935 & 0.001393 \\
	\end{tabular}
	\caption{Times are in seconds. The Graph500 generates the graph instead of loading it into a file. GraphBIG builds the graph and reads in the file simultaneously. These results were averaged across 64 roots. [TODO: Why does the first root often have degree 0?]}
	\label{tab:fileload}
\end{table}

>>>>

\section{Algorithms}

The canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}. The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation. The primary drawback with the Graph500 is it measures a single algorithm: breadth first search (BFS).

This report attempts to add similar rigor to other graph algorithms by borrowing heavily from the Graph500 specification. The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the actual BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}. Some clarifications are in order: for the first kernel, the time to read the graph from disk is not considered. Furthermore, these data structures need only be created once and BFS is performed on 64 roots.

@@

\section{The Existing Systems}
This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}. GraphMat is also included because it has been cited as the most performant \cite{Sundaram:2015:GraphMat}. Our target is shared memory CPU processing.

% Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
% are not considered here because they are a PITA to get working and the authors never responded to my emails.

\section{Experimental Setup}
\subsection{Machine Specifications}
Table~\ref{tab:specs} shows the specifications of the research machine.

@@

\subsection{Datasets}
We use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19, C = 0.19,$ and $D = 1-(A+B+C) = 0.05$.

\section{Perf. Anal.}

\begin{figure}[!htb]
	\includegraphics[width=\columnwidth]{/Users/spollard/Desktop/modern-aRt.pdf}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/bfs_time.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic. [TODO: Replace this with the data with non-zero degree vertices]}
	\label{fig:bfs-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/bfs_dsc.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic. GraphBIG reads in the file and generates the data structure simultaneously so is omitted.}
	\label{fig:bfs-dsc}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/sssp_time.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic.}
	\label{fig:sssp-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/pr_time.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis is logarithmic. GraphMat continues to run until none of the vertices' ranks change. [There is only one data point]}
	\label{fig:pr-time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{graphics/pr_iters.pdf}
	\vspace{-18pt}
	\caption{The $y$-axis \emph{not} logarithmic. GraphMat iterations are measured differently because of the ``think like a vertex'' paradigm and runs unitl none of the vertices' ranks change. [There is only one data point]}
	\label{fig:pr-iters}
\end{figure}

@@

\begin{table}[!htb]
	\centering
	\begin{tabular}{l|c|c|c}
		System & Load Graph & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular} & Run BFS \\ \hline
		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
		GraphBIG & 37.22  & N/A    & 0.1528 \\
		GraphMat & 0.1511 & 1.101  & 0.1031 \\
		GAP      & 2.351  & 0.1935 & 0.001393 \\
	\end{tabular}
	\caption{The above table shows times for $2^{20}$ verties and the times are in seconds. The Graph500 generates the graph instead of loading it into a file. GraphBIG builds the graph and reads in the file simultaneously. These results were averaged across 64 roots.}
	\label{tab:fileload}
\end{table}
